full title

### 2020-10-29 paper/pldi21/combined.tex
Added diagram for combined analysis and more explanation

We define the \textit{combined analysis} of the abstract interpretation and the
sealed symbolic execution.  In the combined analysis, we use the powerset of
symbolic states as the domain $\combdom = \powerset{\symbstset}$, and extend the
view transition as follows:
\[
  \begin{array}{l}
    \combviewtrans{\view}{\view'}(\combelem) =\\
    \qquad \{
      \symbst' \mid \symbst \in S \wedge
      \symbst \symbtrans \symbst' \wedge
      \saconverter(\symbst') = (\view', \_)
    \}\\
    \qquad \cup \left\{
    \begin{array}{ll}
      \{ \asconverter \circ \viewtrans{\view}{\view'}(\bigjoin D) \}
      & \text{if} \; D \neq \varnothing\\
      \varnothing & \text{otherwise}\\
    \end{array}
    \right.
  \end{array}
\]
where $S = \combelem\mid_{\checker}$ and $D =
\dot{\saconverter}(\combelem\mid_{\neg\checker})$.  The dot notation $\dot{f}$
denotes the element-wise extended function of a function $f$.  We could
configure when the sealed symbolic execution or the abstract interpretation
is performed based on the definition of the $\checker$ function and its negation
$\neg\checker$.  It should satisfy the following condition for the soundness and
the termination of the combined analysis:
\[
  \checker(\symbst) \Rightarrow \symbst \symbtrans^k \excst \wedge
  1 < k \leq N
\]
where $N$ is a pre-defined maximum length of the sealed symbolic execution.
The combined analysis just performs one-step symbolic transition for each
symbolic state that passes the filter $\checker$. On the other hand, it converts
remaining symbolic states to the corresponding abstract states, merges them into
a single abstract state, and performs the abstract one-step execution for the
abstract state.

>>>>

\begin{figure}[t]
  \centering
  \includegraphics[width=\linewidth]{img/combined-analysis}
  \vspace*{-2em}
  \caption{The diagram for the extended view transition in the combined analysis.}
  \vspace*{-1em}
  \label{fig:combined-analysis}
\end{figure}

We define the \textit{combined analysis} of the abstract interpretation and the
sealed symbolic execution.  In the combined analysis, we use the powerset of
symbolic states as the domain $\combdom = \powerset{\symbstset}$, and extend the
view transition as follows:
\[
  \begin{array}{l}
    \combviewtrans{\view}{\view'}(\combelem) =\\
    \qquad \{
      \symbst' \mid \symbst \in S \wedge
      \symbst \symbtrans \symbst' \wedge
      \saconverter(\symbst') = (\view', \_)
    \}\\
    \qquad \cup \left\{
    \begin{array}{ll}
      \{ \asconverter \circ \viewtrans{\view}{\view'}(\bigjoin D) \}
      & \text{if} \; D \neq \varnothing\\
      \varnothing & \text{otherwise}\\
    \end{array}
    \right.
  \end{array}
\]
where $S = \combelem\mid_{\checker}$ and $D =
\dot{\saconverter}(\combelem\mid_{\neg\checker})$.  The dot notation $\dot{f}$
denotes the element-wise extended function of a function $f$.  We could
configure when the sealed symbolic execution or the abstract interpretation is
performed based on the definition of the $\checker$ function and its negation
$\neg\checker$.  The combined analysis performs one-step symbolic transition for
each symbolic state that passes the filter $\checker$. On the other hand, it
converts remaining symbolic states to the corresponding abstract states, merges
them into a single abstract state, and performs the abstract one-step execution
for the abstract state.

For example, Figure~\ref{fig:combined-analysis} depicts the
diagram of the view transition from $\view$ to $\view'$ in the combined
analysis.  The element $\combelem$ in $\view$ consists of five symbolic states
from $\nsymbst{1}$ to $\nsymbst{5}$.  The solid box denotes that the corresponding
symbolic state passes the filter $\checker$ and the dotted box denotes that it
does not.  For each symbolic state that passes the filter ($\nsymbst{1}$,
$\nsymbst{2}$, and $\nsymbst{4}$), it applies the symbolic transition $\symbtrans$.
For failed symbolic states ($\nsymbst{3}$ and $\nsymbst{5}$), it applies the
converter $\saconverter$ to convert each of them to the corresponding abstract
state and merges results using the join operator ($\join$).  Then, it utilizes
the abstract semantics via the original view transition
$\viewtrans{\view}{\view'}$ and the converter $\asconverter$ to convert the
result to the corresponding symbolic state.

For the soundness and the termination of the combined analysis, the filter
$\checker$ in the extended view transition should satisfy the following
condition:
\begin{theorem}
  The combined analysis has sound analysis result and terminates in a finite
  time if its abstract semantics is sound and $\checker$ satisfies the following
  condition:
  \[
    \checker(\symbst) \Rightarrow \symbst \symbtrans^k \excst \wedge
    1 < k \leq N
  \]
  where $N$ is a pre-defined maximum length of the sealed symbolic execution.
\end{theorem}
\inred{We prove that any combined analysis defined with the extended view
transition satisfying the above condition always has sound analysis result and
terminates in a finite time. Because of the page limitation, we omit the
proof in this paper and include it in a companion report.}
